Appendix 1: Quantifying Drought

### 2021-04-29 doc/SPEI_appendix.Rmd

```{r}
spei_table <-
  comb %>% 
  group_by(year, dataset) %>% 
  mutate(min = min(spei, na.rm = TRUE)) %>% 
  filter(spei == min) %>% 
  mutate(month = month(yearmonth, label = TRUE)) %>% 
  # summarize(min = min, month = month) %>%
  filter(year %in% c(2002, 2004, 2006, 2009, 1997, 2005, 2010)) %>% 
  select(year, dataset, spei, month) %>% 
  mutate(dataset = str_remove(dataset, "_spei")) %>% 
  pivot_wider(names_from = dataset, values_from = c(spei, month)) %>% 
  add_column(cause = c("ENSO", "ENSO", "ENSO", "AMO", "ENSO", "ENSO", "AMO"), 
             .after = "year") %>% 
  mutate(`Global SPEI` = glue("{round(spei_g, 2)} ({month_g})"),
         `Xavier et al.` = glue("{round(spei_xa, 2)} ({month_xa})"),
         EMBRAPA = glue("{round(spei_st, 2)} ({month_st})"),
         TRMM = glue("{round(spei_trmm, 2)} ({month_trmm})"),
         BDFFP = glue("{round(spei_bdffp, 2)} ({month_bdffp})")) %>% 
  ungroup()
```

>>>>

```{r}
citations <- tribble(~year, ~Reference,
                     2005, "Marengo et al. 2008; Zeng et al. 2008",
                     2010, "Lewis et al. 2011",
                     1997, "McPhaden 1999",
                     2002, NA,
                     2009, NA,
                     2004, NA,
                     2006, NA) 
```

```{r}
spei_table <-
  comb %>% 
  group_by(year, dataset) %>% 
  mutate(min = min(spei, na.rm = TRUE)) %>% 
  filter(spei == min) %>% 
  mutate(month = month(yearmonth, label = TRUE)) %>% 
  # summarize(min = min, month = month) %>%
  filter(year %in% c(2002, 2004, 2006, 2009, 1997, 2005, 2010)) %>% 
  select(year, dataset, spei, month) %>% 
  mutate(dataset = str_remove(dataset, "_spei")) %>% 
  pivot_wider(names_from = dataset, values_from = c(spei, month)) %>% 
  add_column(cause = c("ENSO", "ENSO", "ENSO", "AMO", "ENSO", "ENSO", "AMO"), 
             .after = "year") %>% 
  mutate(`Global SPEI` = glue("{round(spei_g, 2)} ({month_g})"),
         `Xavier et al.` = glue("{round(spei_xa, 2)} ({month_xa})"),
         EMBRAPA = glue("{round(spei_st, 2)} ({month_st})"),
         TRMM = glue("{round(spei_trmm, 2)} ({month_trmm})"),
         BDFFP = glue("{round(spei_bdffp, 2)} ({month_bdffp})")) %>% 
  ungroup()

spei_table <- left_join(spei_table, citations)
```
